---
title: "final project"
author: "LI Dongyang 1155187311"
date: "2023-05-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First, we load data into memory and perform a simple statistical analysis on the data. And "data" stands for the train data, "oosdata" stands for the test data.

```{r}
set.seed(666)
data <- load(url("https://github.com/zhentaoshi/Econ5821/raw/main/data_example/dataset_inf.Rdata"))
oosdata <- load(url("https://github.com/zhentaoshi/Econ5821/raw/main/data_example/data_oos.Rdata"))
summary(cpi)
summary(ppi)
summary(real.cpi)
summary(real.ppi)
```

Then, we join the features, cpi and ppi together using \`month\`.

```{r}

data <- merge(X, cpi, by="month")
data <- merge(data, ppi, by="month")

oosdata <- merge(real.X, real.cpi, by="month")
oosdata <- merge(oosdata, real.ppi, by="month")
```

We first try XGBoost.
As XGBoost using specific data format, we must transform our data to DMatrix.
```{r}
library(xgboost)
library(Matrix)

train_data = data
test_data = oosdata
trainX = data.matrix(train_data[,c(2:152,154)])
trainX = Matrix(trainX,sparse=T)
trainY = train_data[,153]
data = list(x = trainX, y = trainY )
dtrain = xgb.DMatrix(data = data$x, label = data$y)

testX = data.matrix(test_data[,c(2:152,154)])
testX = Matrix(testX,sparse=T)
testY = test_data[,153]
data = list(x = testX, y = testY )
dtest = xgb.DMatrix(data = data$x, label = data$y)

```

After preparing the training data and testing data, we start to train the XGBoost model， we tried different combinations of hyper-parameters. Lastly, we can get the best model below.

```{r}
xgb <- xgboost(data = dtrain,max_depth=4,eta=0.1, nround=253)
#test data (oos r2)
pre_xgb = predict(xgb,newdata = dtest)
rss = var(testY - pre_xgb)
tss = var(testY)
oos_r2 = 1-rss/tss
oos_r2
#train data (is r2)
pre_xgb = predict(xgb,newdata = dtrain)
rss = var(trainY - pre_xgb)
tss = var(trainY)
is_r2 = 1-rss/tss
is_r2
```

Then we use random forest.
```{r}
library(randomForest)

rf = randomForest(CPI ~ . ,ntree=100, data = train_data)
#test data (oos r2)
pre_rf = predict(rf,newdata = test_data)
rss = var(testY - pre_rf)
tss = var(testY)
oos_r2 = 1-rss/tss
oos_r2
#train data (is r2)
pre_rf = predict(rf,newdata = train_data)
rss = var(trainY - pre_rf)
tss = var(trainY)
is_r2 = 1-rss/tss
is_r2
```

kNN
```{r}
library(caret)
control = trainControl(method = 'cv',number = 10)
kNN = train(CPI~.,train_data,
               method = 'knn',
               trControl = control,
               tuneLength = 5)
#test data (oos r2)
pre_kNN = predict(kNN,newdata = test_data)
rss = var(testY - pre_kNN)
tss = var(testY)
oos_r2 = 1-rss/tss
oos_r2
#train data (is r2)
pre_kNN = predict(kNN,newdata = train_data)
rss = var(trainY - pre_kNN)
tss = var(trainY)
is_r2 = 1-rss/tss
is_r2
```

decision tree
```{r}
library(rpart)
DT = rpart(CPI~., train_data)
#test data (oos r2)
pre_DT = predict(DT, newdata = test_data)
rss = var(testY - pre_DT)
tss = var(testY)
oos_r2 = 1-rss/tss
oos_r2
#train data (is r2)
pre_DT = predict(DT, newdata = train_data)
rss = var(trainY - pre_DT)
tss = var(trainY)
is_r2 = 1-rss/tss
is_r2
```

lasso
```{r}
library(glmnet)
x = trainX
y = trainY
cv_lasso = cv.glmnet(x, y)
lasso_result = glmnet(x, y, lambda = cv_lasso$lambda.min)
#test data (oos r2)
pre_lasso = predict(lasso_result, newx = testX)
rss = var(testY - pre_lasso)
tss = var(testY)
oos_r2 = 1-rss/tss
oos_r2
#train data (is r2)
pre_lasso = predict(lasso_result, newx = trainX)
rss = var(trainY - pre_lasso)
tss = var(trainY)
is_r2 = 1-rss/tss
is_r2
```

neural network
```{r}
library(neuralnet)
nn = neuralnet(CPI ~ ., train_data,hidden = 5)
#test data (oos r2)
pre_nn = predict(nn, newdata = test_data)
rss = var(testY - pre_nn)
tss = var(testY)
oos_r2 = 1-rss/tss
oos_r2
#train data (is r2)
pre_nn = predict(nn, newdata = train_data)
rss = var(trainY - pre_nn)
tss = var(trainY)
is_r2 = 1-rss/tss
is_r2
```
By comparing the results of CPI above, we can see that the xgboost has the best r2 in the train data set, but lasso has the best r2 in the test data set.


As for ppi, we first use XGBoost.
```{r}
PPItrainX = data.matrix(train_data[,c(2:153)])
PPItrainX = Matrix(PPItrainX,sparse=T)
PPItrainY = train_data[,154]
data = list(x = PPItrainX, y = PPItrainY )
PPIdtrain = xgb.DMatrix(data = data$x, label = data$y)

PPItestX = data.matrix(test_data[,c(2:153)])
PPItestX = Matrix(PPItestX,sparse=T)
PPItestY = test_data[,154]
data = list(x = PPItestX, y = PPItestY )
PPIdtest = xgb.DMatrix(data = data$x, label = data$y)
```

```{r}
PPIxgb <- xgboost(data = PPIdtrain, max_depth = 12, eta = 0.3,nround=60)
#test data (oos r2)
pre_PPIxgb = predict(PPIxgb,newdata = PPIdtest)
rss = var(PPItestY - pre_PPIxgb)
tss = var(PPItestY)
oos_r2 = 1-rss/tss
oos_r2
#train data (is r2)
pre_PPIxgb = predict(PPIxgb,newdata = PPIdtrain)
rss = var(PPItrainY - pre_PPIxgb)
tss = var(PPItrainY)
is_r2 = 1-rss/tss
is_r2
```

randomForest
```{r}
rf = randomForest(PPI ~ . , data = train_data,ntree = 500)

#test data (oos r2)
pre_rf = predict(rf,newdata = test_data)
rss = var(PPItestY - pre_rf)
tss = var(PPItestY)
oos_r2 = 1-rss/tss
oos_r2
#train data (is r2)
pre_rf = predict(rf,newdata = train_data)
rss = var(PPItrainY - pre_rf)
tss = var(PPItrainY)
is_r2 = 1-rss/tss
is_r2
```

kNN
```{r}
control = trainControl(method = 'cv',number = 10)
kNN = train(PPI~.,train_data,
               method = 'knn',
               trControl = control,
               tuneLength = 5)
#test data (oos r2)
pre_kNN = predict(kNN,newdata = test_data)
rss = var(PPItestY - pre_kNN)
tss = var(PPItestY)
oos_r2 = 1-rss/tss
oos_r2
#train data (is r2)
pre_kNN = predict(kNN,newdata = train_data)
rss = var(PPItrainY - pre_kNN)
tss = var(PPItrainY)
is_r2 = 1-rss/tss
is_r2
```

decision tree
```{r}
DT = rpart(PPI~., train_data)
#test data (oos r2)
pre_DT = predict(DT, newdata = test_data)
rss = var(PPItestY - pre_DT)
tss = var(PPItestY)
oos_r2 = 1-rss/tss
oos_r2
#train data (is r2)
pre_DT = predict(DT, newdata = train_data)
rss = var(PPItrainY - pre_DT)
tss = var(PPItrainY)
is_r2 = 1-rss/tss
is_r2
```

lasso
```{r}
x = PPItrainX
y = PPItrainY
cv_lasso = cv.glmnet(x, y)
lasso_result = glmnet(x, y, lambda = cv_lasso$lambda.min)
#test data (oos r2)
pre_lasso = predict(lasso_result, newx = PPItestX)
rss = var(PPItestY - pre_lasso)
tss = var(PPItestY)
oos_r2 = 1-rss/tss
oos_r2
#train data (is r2)
pre_lasso = predict(lasso_result, newx = PPItrainX)
rss = var(PPItrainY - pre_lasso)
tss = var(PPItrainY)
is_r2 = 1-rss/tss
is_r2
```

neural network
```{r}
nn = neuralnet(PPI ~ ., train_data,hidden = 20)
#test data (oos r2)
pre_nn = predict(nn, newdata = test_data)
rss = var(PPItestY - pre_nn)
tss = var(PPItestY)
oos_r2 = 1-rss/tss
oos_r2
#train data (is r2)
pre_nn = predict(nn, newdata = train_data)
rss = var(PPItrainY - pre_nn)
tss = var(PPItrainY)
is_r2 = 1-rss/tss
is_r2
```

By comparing the results of PPI above, we can see that the xgboost has the best r2 in the train data set, but lasso has the best r2 in the test data set.

We can now use lasso to make the prediction of both CPI and PPI.